\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-20 paper/paper.tex
Updates to paper.

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions - with innovations targeted at taking
advantage of growing flash speeds. Unfortunately their performance is hampered
by the need to maintain and replicate a write ahead log to guarantee
availability in the face of machine and storage failures. Cyclone offers a
solution to this problem by leveraging a small amount of directly attached
non-volatile memory (NVM) such as NVDIMMs to maintain a two level log - the upper
level in a small amount of NVM draining into a larger lower level log
maintained on flash. Replicating the smaller log kept in NVMs transforms the
replication problem into a packet switching problem - the leader sends the exact
same packet to all follower replicas and we solve the problem as such by using
the RAFT protocol implemented as a software multicast module on the Data Plane
Development Kit (DPDK). Second, we levergage the observation that key value store
interfaces are commutative for operations to different keys to scale our two
level log horizontally into a number of physical two level logs - with a hash of
the key being used to choose the right log. Each log is replicated by an
independent instance of RAFT, thereby leveraging the scalability afforded by
classic software packet switching when using multiple CPU cores. Cyclone is able
to replicate millions of small updates a second using only commodity 10 gigabit
ethernet adapters and improves the performance of RocksDB - a popular persistent
key value store by orders of magnitude when compared to its own write ahead log
without replication.
\end{abstract}

@@

We implement the common case flow in RAFT as a software packet switch using
DPDK. The dataplane now encompasses the NIC and the CPU caches with the NIC
placing packet data directly in the CPU cache via DDIO~\cite{ddio} - a standard
feature of most NICs today. The dataplane for the RAFT protocol is therefore
separated from the control plane. The control plane now runs on a CPU core and
takes care of checking adherence to the protocol: checking for correct term, log
index and that the replica receiving the packet from the client is indeed the
leader. In addition, the control plane is responsible for non common-case
activities such as dealing with timeout and leader elections.

@@

\section{Horizontal Scaling}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated core. Further, as we shortly describe we
assume a concurrent key value store where a set of CPU cores independently run
operations from the physical logs on the shared memory key-value store data
structure, synchronizing as necessary.

We begin by describing how a key-value store API can be mapped on these
independent physical logs without divergence in replica state. We then discuss
how different pieces of work related to replication and the key value store
itself are distributed across available CPU cores and available NIC queues.

@@

Further relaxations to the mapping scheme are possible when operations to the
same key are also commutative. A good example is when key value stores are used
to maintain counters. Different increment operations to the same counter are
commutative if one is prepared to tolerate non-determinism in the sequence of
updates seen by a reader with the counter reaching the same eventual value. This
specific use-case, in fact, lead to the addition of the merge operation in
Rocksdb, a key value store that we use for evaluation. Merge operations
therefore could be dispatched to a random physical log with no harm. For this
paper however, we stick on the stricter requirement that operations to the same
key are not commutative.

\subsection{Core Assignment}

TBD: mention no linearizability for reads at the moment

\subsection{NIC Queue Assignmnt}

>>>>

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions - with innovations targeted at taking
advantage of growing flash speeds. Unfortunately their performance is hampered
by the need to maintain and replicate a write ahead log to guarantee
availability in the face of machine and storage failures. Cyclone offers a
solution to this problem by leveraging a small amount of directly attached
non-volatile memory (NVM) such as NVDIMMs to maintain a two level log - the upper
level in a small amount of NVM draining into a larger lower level log
maintained on flash. Replicating the smaller log kept in NVMs transforms the
replication problem into a packet switching problem - the leader sends the exact
same packet to all follower replicas and we solve the problem as such by using
the RAFT protocol implemented as a software multicast module on the Data Plane
Development Kit (DPDK). Second, we levergage the observation that key value store
interfaces are commutative for operations to different keys to scale our two
level log horizontally into a number of physical two level logs - with a hash of
the key being used to choose the right log. Each log is replicated by an
independent instance of RAFT, thereby leveraging the scalability afforded by
classic software packet switching when using multiple threads. Cyclone is able
to replicate millions of small updates a second using only commodity 10 gigabit
ethernet adapters and improves the performance of RocksDB - a popular persistent
key value store by orders of magnitude when compared to its own write ahead log
without replication.
\end{abstract}

@@

We implement the common case flow in RAFT as a software packet switch using
DPDK. The dataplane now encompasses the NIC and the CPU caches with the NIC
placing packet data directly in the CPU cache via DDIO~\cite{ddio} - a standard
feature of most NICs today. The dataplane for the RAFT protocol is therefore
separated from the control plane. The control plane now runs in a thread pinned
to a dedicated CPU core and takes care of checking adherence to the protocol:
checking for correct term, log index and that the replica receiving the packet
from the client is indeed the leader. In addition, the control plane is
responsible for non common-case activities such as dealing with timeout and
leader elections.

@@

\section{Horizontal Scaling}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated CPU core. Further, we assume a concurrent
key value store where a set of CPU cores independently run operations from the
physical logs on the shared memory key-value store data structure, synchronizing
as necessary.

We begin by showing how we divide CPU cores and NIC resources on the plaform across
replication and key value store work, and show how a KV store request flows
across them.  We then describe how a key-value store API can be mapped to
independent physical logs without divergence in replica state. Finally, we
describe special handling in Cyclone for multi-key operations.

\subsection{Core Allocation}
A key-value store integrated with Cyclone runs two types of threads. One type of
thread does key-value store work. This includes servicing lookup and update
requests and background threads that flush in-memory data structures to
secondary storage and do maintainence work on secondary storage data structures
such as compacting log-structured merge trees~\cite{lsmtree}. We use a dedicated
thread per-physical log to run the corresponding control plane including the
associated RAFT state machine. We use static assignment of threads to cores as
recommended for packet switching libraries such as DPDK for the best possible
performance. Therefore we use the term threads and cores interchangeably in this
paper. Client requests are always received by the control plane core for
the physical log the request is directed to. A response for the core is always
sent by the KV store core that finally handles the request.

An important resource for Cyclone is the NIC that must be shared among the
control plane cores that receive and replicate requests and the KV store cores
that send back responses. Most NICs provide independent transmit/receive queue
pairs and we aim for synchronization free access to these to aid horizontal
scalability with respect to the network resource. DPDK already provides fairly
scalable access to NIC queues, with the queue number being a parameter to send
and receive calls. All that remains for us is to map the queues to the control
plane and KV store cores, keeping in mind that there may be multiple NICs on the
system. We do this by iterating over all control plane cores followed by all KV
store cores. We assign two queue pairs to each control plane core, one to
receive requests from the client and the other transmit and receive replication
protocol related messages. We assign one queue pair to each KV store core to
send back responses to clients. We are careful to spread the different queue
pairs for a core across the available NICs as far as possible. This is important
due to the asymmetry between the different types of traffic. Queue pairs
carrying replication traffic will likely be more busy that those carrying client
request responses due to the amplification to multicast the same message to
different replicas.

The flow for an update operation in Cyclone is as follows. The network packet is
received by the control plane for the targeted physical log. The packet is then
immediately sent out for replication as described in the previous section. When
responses from a majority quorum is received, the request is passed to the
targeted KV store core for execution. The KV store core process the request and
sends back the response to the client. We note that replication is pipelined and
decoupled from execution and thus multiple updates requests are usually in
flight for replication at a time.

The flow for a read operation is simpler. It is still targeted to a physical log
but instead to replicating it the control plane core simply passes it on
straight away to the KV store core for execution and response.

The control plane cores and KV store cores communicate through a one-way
lock-free FIFO queue passing pointers to packets. We leverage the convenient
ring library in DPDK for this.

Finally an asynchronous source of work in the system is maintenance of the
two-level log. We depart from a 1:1 mapping between the first level NVM log and
the second level flashlog here. Each physical log consists of one NVM log
managed by the corresponding control plane core and multiple flashlogs each of
which is managed as an additional responsibility by a KV store core. We
therefore enforce a many to one mapping between key value store cores and
control plane cores - thereby partitioning KV store cores across control plane
cores. A request picks a control plane core and any one of the associated KV
store cores for execution. The request is logged in the NVM log of the control
plane core and garbage collected once it is drained to the flashlog of the
selected KV store core.

@@

Further relaxations to the mapping scheme are possible when operations to the
same key are also commutative. A good example is when key value stores are used
to maintain counters. Different increment operations to the same counter are
commutative if one is prepared to tolerate non-determinism in the sequence of
updates seen by a reader with the counter reaching the same eventual value. This
specific use-case, in fact, lead to the addition of the merge operation in
Rocksdb, a key value store that we use for evaluation. Merge operations
therefore could be dispatched to a random physical log with no harm. For this
paper however, we stick to the stricter requirement that operations to the same
key are not commutative. However, Cyclone's client side libraries leave it to
the user to specify the exact physical log and key-value thread to use to enable
more relaxed mappings when necessary.

We note that commutativity as assumed by Cyclone is an \emph{API level
  property}. There are no assumption as to how application cores synchronize to
access the shared key-value state or the underlying data structures for holding
key-value pairs.

Finally, Cyclone provides seraializable reads but does not - at the moment -
guarantee the ability to read one's own writes, repeatable reads or more
generally, linearizable reads. The reason is that on a failover the new leader
replica might still have update operations in its NVM logs that are yet to be
applied to the key value store. However the original failed leader replica might
have applied these updates and a client might have already observed them. Users
who desire linearizable reads can issue a special API call to ensure that reads
are also routed through the logs. Flushing the NVM logs on failover is a better
solution that we are currently in the process of engineering into Cyclone.
